Prediction Developing Countries' Life Expectancy

### 2025-02-21 paper.qmd
fixed titles

@@

library(tidyverse)
library(arrow)
library(knitr)
library(modelsummary)
library(janitor)
library(ggplot2)
library(knitr)
```

@@

alcohol_status <- data_interest %>%
  group_by(alcohol) %>%
  summarise(Count = n()) %>%
  as.data.frame()

# Display the table using kable
# kable(alcohol_status, caption = "Number of Countries with Negligible and Consumed Alcohol Data")

@@

## Basic Linear Model Summary

Model is Life Expectancy = GDP Per Capita + Infant Deaths (per 1000 people) + HIV/AID Deaths (per 1000 live births) + Health Expenditure (as a percentage of GDP per Capita) + Average Number of Years of Schooling + Alcohol Consumption (Negligible Consumption vs. Alcohol is Consumed)

```{r}
#| warning: false
#| message: false
#| echo: false

preliminary_model = lm(life_expectancy ~ infant_deaths + hiv_aids + gdp_capita + 
                         schooling + percentage_expenditure + alcohol, 
                       data=data_interest)

#summary(preliminary_model)
modelsummary(preliminary_model)

@@

plot(preliminary_model, which = 1)
plot(preliminary_model, which = 2)
standard_residuals = rstandard(preliminary_model)
plot(data_interest$gdp_capita, standard_residuals, main="Standardized Residuals V. GDP Per Capita", 
     xlab="GDP Per Capita", ylab="Standardized Residuals")

@@

### Linear Regression Residual Assumptions:

#### Linearity
#### Normal Errors
#### Constant Variance Assumption

# Conclusion

>>>>

# Contributions

@@

library(tidyverse)
library(knitr)
library(janitor)
library(ggplot2)
```

@@

@@

```{r}
#| label: tbl-modelresults
#| warning: false
#| message: false
#| echo: false
preliminary_model = lm(life_expectancy ~ infant_deaths + hiv_aids + gdp_capita + 
                         schooling + percentage_expenditure + alcohol, 
                       data=data_interest)

#summary(preliminary_model)

model_table <- data.frame(
  Variable = c("(Intercept)", "Infant Deaths", "HIV/AIDS", "GDP per Capita", 
               "Schooling", "Percentage Expenditure", "Alcohol (Negligible)"),
  Estimate = c(49.50, -0.002131, -0.6713, 7.542e-06, 1.701, 0.00109, -0.670),
  Std_Error = c(0.7476, 0.0009369, 0.01917, 5.514e-05, 0.06129, 0.0003605, 0.2859),
  T_Value = c(66.216, -2.274, -35.018, 0.137, 27.757, 3.025, -2.343),
  P_Value = c("< 2e-16", "0.02309", "< 2e-16", "0.89121", "< 2e-16", "0.00253", "0.01924")
)

# Generate a nicely formatted table
kable(model_table, digits = 4, align = "c",
      col.names = c("Variable", "Estimate", "Std. Error", "t Value", "Pr(>|t|)"))

## Confouding Plots

```{r}
#| label: fig-confoundingplots
#| layout-ncol: 2
#| fig-cap: ""
#| fig-subcap: ["", ""]
#| fig-align: "center"
#| warning: false
#| message: false
#| echo: false
#| fig.show: "hold"

ggplot(data_interest, aes(x = percentage_expenditure, y = gdp_capita)) +
  geom_point(alpha = 0.6) +
  geom_smooth() +
  theme_minimal() +
  labs(title = "Health Expenditure can Predict GDP Per Capita",
       x = "Health Expenditure (as a percentage of GDP per capita)", 
       y = "GDP per Capita")

ggplot(data_interest, aes(x = percentage_expenditure, fill = alcohol)) +
  geom_histogram(binwidth = 250, color = "black", alpha = 0.6, position = "identity") +
  theme_minimal() +
  labs(title = "Countries that Consume Alcohol Spend More on Health",
       x = "Health Expenditure (as a percentage of GDP per capita)",
       y = "Count",
       fill = "Alcohol Consumption") +
  scale_fill_manual(values = c("Negligible" = "skyblue", "Consumed" = "salmon"))

@@

plot(preliminary_model, which = 1)
plot(preliminary_model, which = 2)
standard_residuals = rstandard(preliminary_model)
plot(data_interest$schooling, standard_residuals, main="Standardized Residuals V. Schooling", 
     xlab="Average Years in School", ylab="Standardized Residuals")

@@
